# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50.
```{r}
for (i in 1:50) {
  print(i) #print current integer
}

```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
i <- 0 #provide the starting value for i
sum <- 0 #provide the starting value for sum
while (i < 1000) {
  i <- i+1 #add 1 to get the next integer
  sum <- sum + i #add this value to the previous sum
}
sum #print the sum
```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
i <- 0 #provide the starting value for i
sum <- 0 #provide the starting value for sum
while(i < 1000) {
  i <- i+2 #add 2 to get the next even integer
  sum <- sum + i #add this value to the previous sum
}
sum #print the sum
```


C. Now, repeat A and B WITHOUT using a loop.
```{r}
sum(0:1000) #sum of all integers from 0 to 1000

nums <- 0:1000 #initialize a vector of integers from 0 to 1000
sum(nums[lapply(nums, "%%", 2) == 0]) #find the even numbers from nums and sum all of them up
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.

```{r}
survey.clean <- survey #initialize the survey.clean dataframe
for (row in 1:nrow(survey.clean)) {
  for(col in 1:ncol(survey.clean)){
    if (survey.clean[row, col] >= 1 & survey.clean[row, col] <= 5){
      next #skip the values that are correct
    } else {
      survey.clean[row, col] <- NA #replace invalid values with NA
    }
  }
}
survey.clean #print out cleaned survey
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
survey.clean$invalid.answers <- NA #initialize column of NAs
for (i in 1:nrow(survey.clean)){ #repeat over the different rows
  y <- survey.clean[i,1:ncol(survey.clean)-1] #involve all columns except for final one that has the NAs
  x <- is.na(y) #check if cell is NA
  survey.clean$invalid.answers[i] <- sum(x) #add number of NAs
}
survey.clean
```