# pre-class
Kevin Chen

Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f1("cat", "c")


f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

f3 <- function(x, y) {
  rep(y, length.out = length(x))
}

f3(c(2, 2, 1, 13, 1, 2, 23, 15, 1), c(1, 2, 3, 4, 5, 6))
```

f1 checks if a string has a given prefix. A better name for f1 would be `check_prefix`.
      
f2 drops the last element of x. If the length of x is less than or equal to 1, f2 returns NULL. A better name for f2 would be `drop_last`.
       
f3 repeats each element of y for every element of x. If x is longer than y, f3 will begin to recycle the elements of y. A better name of f3 would be `recycle_second`.

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

```{r}
?rnorm #check documentation for rnorm()
?MASS::mvrnorm #check documentation for mvnorm()

```
`rnorm()` generates random samples from a univariate normal distribution, whereas `mvrnorm()` generates samples from multivariate normal distribution. `rnorm()` takes the inputs of n (number of observations), mean (vector of means), and sd (vector of standard deviations), whereas `mvrnorm()` takes the inputs of n (number of samples required), mu (a vector giving the means of the variables), and Sigma (a symmetric matrix that specifies the covariance matrix of the variables). To make the two functions consistent, I would modify the names of the inputs of `mvrnorm()` to make them to be the same of the names of the inputs of `rnorm()`. I would change mu to mean and Sigma to sd. However, sd may not make sense in the context of `mvrnorm()` because `mvrnorm()` takes a covariance matrix instead of a vector of standard deviations. Overall, though, both functions are consistent on their own, as both follow the same input naming convention throughout.


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset.
```{r}
library(dplyr)
head(mtcars) #check mtcars dataset
lapply(mtcars, function(x) sd(x)/mean(x)) #lapply with the anonymous function
```

4. Use vapply() to:
  a. Compute the standard deviation of every column in a numeric data frame.
```{r}
class(mtcars) #check if mtcars is a numeric dataframe
#use mtcars as example dataframe for this exercise
vapply(mtcars, sd, numeric(1)) #calculate sd for every column

```
    
  b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)

```{r}
library(help = "datasets") #look for built-in datasets in R
head(ChickWeight)
class(ChickWeight) #check class of ChickWeight

vapply(ChickWeight[vapply(ChickWeight, is.numeric, logical(1))], sd, numeric(1)) #calculate sd for every numeric column of ChickWeight
```
